ICT and Future Productivity:\\Evidence and Theory of a GPT\thanks{Correspondence: Department of Economics, Boston College, 140 Commonwealth Avenue, Chestnut Hill, MA 02467. Email: brianti@bc.edu (Marco Brianti) and gati@bc.edu (Laura G\'ati).}

### 2018-07-12 Latex/Draft_Summer2018/Summer2018.tex
editing the draft

As a first-pass test, % cite Barsky Basu Lee and Forni or whatevs
we check if our results are robust once we remove all the forward looking variables whose fluctuations may be unrelated to sector-specific technological changes: consumption and output. Technically speaking, consumption and output may Granger-cause future TFP for reasons which are orthogonal to ICT shocks. For example, the forward-looking nature of consumption may provide the VAR with information regarding future changes in TFP not related to an ICT shock, which our identification strategy is not able to filter out. Our second test is running a VAR in which we identify both ICT shocks and news shocks. In particular, we apply a sequential identification where we first identify a news shock in the spirit of \cite{barsky2011news}, and subsequently we identify our sectoral ICT shock using our original identification strategy. This second strategy has the specific purpose of filtering out all the current movements in forward-looking variables related to future fluctuations of TFP which are not related to current ICT shocks, as such fluctuations are all captured by the identified news shock.

@@

In this part of the paper, we employ a first-pass test where we attempt to show that our ICT shock is not confounded with future signals of TFP unrelated to the ICT sector.

As discussed by \cite{sims2012news}, \cite{forni2014sufficient}, and \cite{barsky2015whither} the presence of forward-looking variables in the VAR is a necessary condition to correctly identify a news shock. In particular, since the univariate TFP process cannot predict future TFP,\footnote{It is well-accepted that TFP growth can be approximated as a white noise.} a news shock can be identified if there are variables that granger-cause TFP inside the VAR. For example, \cite{beaudry2006stock} use current movements in stock prices to predict future productivity, but also contemporary fluctuations in consumption, hours, investment, may implicitly reflect information regarding future technical change.

Thus, as a simple test we propose to redo the exercise described in Section \ref{section:mainSpec} without two forward-looking macroeconomic variables not directly related with the ICT sector: consumption and output.\footnote{Notice that output implicitly reflect the behavior of hours and investment once we control for consumption.} If we are actually confounding news shocks with our ICT shock series then we would expect to see some differences in this second estimate. In other words, once our VAR is deprived by a source of information useful to recover a news shock, then the confounded part should either decreases or cancels out.

As hinted above, in this section we estimate the following 3-dimensional system
\begin{equation}\label{eq:noForVariables}
\begin{pmatrix}
TFP_t \\ 
ICTI_t \\
RP_t \\
\end{pmatrix} = B(L) \begin{pmatrix}
TFP_{t-1} \\ 
ICTI_{t-1} \\
RP_{t-1} \\
\end{pmatrix} + i_t
\end{equation}
where  $TFP_t$, $ICTI_{t}$, and $RP_t$ have the same meaning and transformation as in System \ref{eq:mainSpecification}. As before, dataset is quarterly and covers the U.S. economy from 1989:Q1 to 2017:Q1. Finally, according to the Bayesian Information Criterion (BIC), we choose one lag.\footnote{The Hannan-Quinn Criterion (HQ) suggests to use two lags. Results remains consistent following this second criterion.} Using the same identification strategy presented in \ref{section:empiricalstrategy_simple}, we recover a series of ICT shocks which is correlated at $95$\% with the one estimated in 5-dimensional system. In particular, in Appendix \ref{section:removing_FL_Var}, Figure \ref{fig:shockSeries_removing} shows ICT shock series for both the 3-dimensional system and 5-dimensional one. In line with the correlation coefficient, the two series almost overlap each others suggesting that our initial result is not driven by the forward-looking power of consumption and output on TFP.

However, although this result is encouraging, this first-pass test has obviously a weak power because the jump of ICTI on impact may be triggered by future exogenous signal of TFP. In this case, ICTI would be itself the source to granger-cause TFP and our ICT shock would be upward biased by news. To amend this deeper concern, we then need to employ a more complicated strategy presented in the next section.

>>>>

As a first-pass test, % cite Barsky Basu Lee and Forni or whatevs
we check if our results are robust once we remove all the forward-looking variables whose fluctuations may be unrelated to sector-specific technological changes: consumption and output. Technically speaking, consumption and output may Granger-cause movements in future TFP for reasons which are orthogonal to ICT shocks. For example, the forward-looking nature of consumption may provide the VAR with information regarding future changes in TFP not related to an ICT shock, which our identification strategy is not able to filter out.

Our second test is running a VAR in which we identify both ICT shocks and news shocks. In particular, we apply a sequential identification where we first identify a news shock in the spirit of \cite{barsky2011news}, and subsequently we identify our sectoral ICT shock using our original identification strategy. This second strategy has the specific purpose of filtering out all the current movements in forward-looking variables related to future fluctuations of TFP which are not related to current ICT shocks, as such fluctuations are all captured by the identified news shock.

@@

In this part of the paper, we employ a first-pass test of whether our ICT shock may be confounded with future signals of TFP unrelated to the ICT sector.

As discussed by \cite{sims2012news}, \cite{forni2014sufficient}, and \cite{barsky2015whither} the presence of forward-looking variables in the VAR is a necessary condition to correctly identify a news shock. In particular, since the univariate TFP process cannot predict its own future values,\footnote{It is well-accepted that TFP growth can be approximated as a white noise process, or, equivalently, that the TFP level process is a random walk.} a news shock can only be identified if there are variables that Granger-cause TFP inside the VAR. For example, \cite{beaudry2006stock} use current movements in stock prices to predict future productivity, but also contemporary fluctuations in consumption, hours, investment, may implicitly reflect information regarding future technical change.

Thus, as a simple test we redo the exercise described in Section \ref{section:mainSpec}, omitting the two forward-looking macroeconomic variables not directly related with the ICT sector: consumption and output.\footnote{Notice that output implicitly reflects the behavior of hours and investment once we control for consumption.} If we are actually confounding news shocks with our ICT shock series then we would expect to see some differences in this second estimate, because absent the two forward-looking variables, our VAR is deprived of a source of information useful to recover a news shock and should thus no longer be able to capture it appropriately.

Without consumption and output, we thus estimate the following three-dimensional system
\begin{equation}\label{eq:noForVariables}
\begin{pmatrix}
TFP_t \\ 
ICTI_t \\
RP_t \\
\end{pmatrix} = B(L) \begin{pmatrix}
TFP_{t-1} \\ 
ICTI_{t-1} \\
RP_{t-1} \\
\end{pmatrix} + i_t
\end{equation}
where  $TFP_t$, $ICTI_{t}$, and $RP_t$ are defined and treated as in System \ref{eq:mainSpecification}. As before, the dataset is quarterly and covers the U.S. economy from 1989:Q1 to 2017:Q1. Finally, analogously to before, we follow the Bayesian Information Criterion (BIC) to choose one lag. Using the same identification strategy presented in \ref{section:empiricalstrategy_simple}, we recover a series of ICT shocks. The newly recovered series is correlated at $95$\% with the one estimated in five-dimensional system, indicating that our main specification is not contaminated by news. Figure \ref{fig:shockSeries_removing} in Appendix \ref{section:removing_FL_Var} shows the ICT shock series for both the three-dimensional system and the five-dimensional one. In line with the very high correlation coefficient, the two series overlap perfectly, suggesting that both procedures recover the same structural shock series. Thus, our first-pass test indicates that our initial result is not driven by the forward-looking power of consumption and output on TFP.

However, although this result is encouraging, the first-pass test has weak power because the jump of ICTI on impact may itself be triggered by exogenous signals on future TFP. In this case, ICTI would be itself the source to Granger-cause TFP and our ICT shock would be upward biased by news. To address this deeper concern, we proceed to employ a more complicated strategy aimed at directly capturing, and filtering out news.
